---
title: "Project Regression Models"
autor: "KS"
date: "Friday, December 19, 2014"
output: pdf_document
---

#Executive Summary

Working in the Motor Trend industry raises many questions. Here we are looking at manual cars versus automatics. Looking at a data set of a collection of cars(mtcars), we are interested in exploring the relationship between a set of variables and MPG (outcome). In this document of particualr interest are the questions:

- “Is an automatic or manual transmission better for MPG”
- "Quantify the MPG difference between automatic and manual transmissions"

In our analysis we can see that the mpg is greater for manual comapared with automatic motor vehicles. 

#Details

```{r echo=FALSE, results="hide"}
d <- as.data.frame(mtcars)  # get my data

# explore my data
dim(d) 
head(d)
summary(d)

# Sort my data for manuipulation
y_mpg<-d$mpg
x_man_aut<-d$am  # automatic = 0# manaual =1
y_mpg
x_man_aut

## some more data manipulation 
dat<-data.frame(x_man_aut,y_mpg) # make a dataframe
# add a column based on a critereia
dat$ManAut <- ifelse(dat$x_man_aut ==0, "Automatic", "Manual")
# make a gglot from this 

##dat_man<-subset(dat,ManAut=="Manual") another way of doing this 
dat_man<-dat[dat$ManAut=="Manual",]
dat_aut<-dat[dat$ManAut=="Automatic",]
dim(dat)

dim(dat_man)
dim(dat_aut)

## look at the means for automatic and manual
mean(dat_man$y_mpg)
mean(dat_aut$y_mpg)

### Sort the max at top 
dat_man<-dat_man[with(dat_man, order(-y_mpg)),]
dat_aut<-dat_aut[with(dat_aut, order(-y_mpg)),]


## take the top 5 for one model of the dfs
t5_man<-dat_man[1:5,]
t5_aut<-dat_aut[1:5,]


## combine these dfs
t5_manaut<-rbind(t5_man,t5_aut)

## TRY THIS PLOT ++  this can be a red model of the top 5 for aut and man 


### GET THE BOTTOM 5 
b5_man<-dat_man[(nrow(dat_man)-5):nrow(dat_man),]
b5_aut<-dat_aut[(nrow(dat_aut)-5):nrow(dat_aut),]


## combine these dfs
b5_manaut<-rbind(b5_man,b5_aut)


### RE-Sort so the min is at the top for the next model 
dat_man<-dat_man[with(dat_man, order(-y_mpg)),]
dat_aut<-dat_aut[with(dat_aut, order(-y_mpg)),]


## add an n column to my dfs
dat_aut$n<-1:nrow(dat_aut)
dat_man$n<-1:nrow(dat_man)


## TRY THIS PLOT ++  this can be a blue model of the bottom 5 for aut and man 


### BUILD My MODELS
fit<-lm(y_mpg~x_man_aut)
## for top 5
fit_t5<-lm(t5_manaut$y_mpg~t5_manaut$x_man_aut)
## for bottom 5
fit_b5<-lm(b5_manaut$y_mpg~b5_manaut$x_man_aut)


### Build more models 

fit_man<-lm(dat_man$y_mpg~dat_man$n)
fit_aut<-lm(dat_aut$y_mpg~dat_aut$n)


## look at the coefficients 

## look at the summarys for comparison
summary(fit)
summary(fit_t5)
summary(fit_b5)

## look at the anovas for comparison 
anova(fit)
anova(fit_t5)
anova(fit_b5)

```


Here I will look at the plots of manual cars versus automatics for miles per gallon.


```{r echo=F}
### PUT MY PLOT CODE IN HERE THAT I WANT TO BE VISIBLE 

par(mfrow=c(1,2)) # 1 row 2 columns of plots 

################### DO PLOT 1 

plot(dat$x_man_aut,dat$y_mpg,pch=21,col="black",bg="black",main="MPG V Automatic/Manual(T5/B5)",xlab="Transmission(0=automatic,1=manual)",ylab="Miles/(US) gallon[mpg]") # this will already have been plotted 
points(t5_manaut$x_man_aut,t5_manaut$y_mpg ,pch=21,col=((t5_manaut$x_man_aut=="1" | t5_manaut$x_man_aut=="0")*1+1),bg="black")
points(b5_manaut$x_man_aut,b5_manaut$y_mpg ,pch=21,col=((b5_manaut$x_man_aut=="1" | b5_manaut$x_man_aut=="0")*1+2),bg="black")
## abline for them all 
abline(fit,col="black",lwd=2) # this is one model 
## for top 5
abline(fit_t5,col="red",lwd=2) # this is one model 
## for bottom 5
abline(fit_b5,col="green",lwd=2) # this is one model 

## Add legend 
legend("topleft",lty=1 , pch=21,col = c("red","black", "green"), legend = c("Top 5", "All","Bottom 5"))

##################### DO PLOT 2 

## do another plot 
plot(jitter(dat_aut$n),dat_aut$y_mpg,pch=19,ylim=c(min(dat_aut$y_mpg),max(dat_man$y_mpg)),main="MPG V Automatic/Manual",xlab="n=1...19",ylab="Miles/(US) gallon[mpg]")  ##use jitter to show >1 points
points(jitter(dat_man$n),dat_man$y_mpg,pch=19,col="red")

## add the models to the plots 
abline(fit_man,col="red")
abline(fit_aut)
## Add legend 
legend("topright",lty=1 , pch=19,col = c("red","black"), legend = c("Manual", "Automatic"))

```
**Figure 1:**

MPG V Automatic/Manual (Top5/Bottom5): A look at the miles per gallon per automatic/manaul transmission motor vehicle. The top 5 and bottom 5 are also looked at here.

MPG V Automatic/Manual: An alternative look at the miles per gallon per automatic/manaul transmission motor vehicle.   


```{r echo=FALSE, results="hide"}
## Here we look at the coefficients 

## look at the summarys for comparison
summary(fit)$coeff
summary(fit_t5)$coeff
summary(fit_b5)$coeff

## look at the anovas for comparison 
anova(fit)
anova(fit_t5)
anova(fit_b5)

```


From the MPG V Automatic/Manual graph in Figure 1 we can look at the coefficients of the 2 models 

```{r}

## look at the summarys for comparison
summary(fit_man)$coeff
summary(fit_aut)$coeff

## look at the anovas for comparison 
anova(fit_man)
anova(fit_aut)

```

From this we can say: 

- manual appears to give better mpg compare with automatic 
- a note on the the p-values 
- a note a confidence intervals 
- a note on the different models 
- other notes 


**Note: The code for this entire documaent can be found [here](https://github.com/HattrickNZ/regressionModels) **


#Appendix 

```{r echo=FALSE}
library(datasets); data(mtcars); require(stats); require(graphics)
pairs(mtcars, panel = panel.smooth, main = "mtcars data", col = 3)
```

**Figure A1:** A look at all the mtcars data.





```{r echo=FALSE, results="hide"}
# TO DELETE
## this is how the project will be greaded!!!


#1. Did you interpret your coefficients correctly?

#2. Did you do some exploratory data analyses?

#3. Did you fit multiple models and detail your strategy for model selection?

#4. Did you answer the questions of interest or detail why the question(s) 
#is (are) not answerable?

#5. Did you do a residual plot and some diagnostics?

#6. Did you quantify the uncertainty in your conclusions and/or perform an inference correctly?

#7. Did you communicate your findings?

#8. Was the report brief (about 2 pages long)?

#9. Did it include an executive summary? see above 

#10. Was the report done in Rmd (knitr)?

```
